\documentclass{subfiles}
\begin{document}
\chapter{Results}\label{chap:4}
In this chapter, we present the results of our parameter optimization and dynamical simulations aimed at validating and operationalizing the Morse double-well system. Our primary objectives are twofold: first, to determine optimal parameters for two distinct configurations ($C_I$ and $C_{II}$) of the Morse double-well that exhibit the desired energy and entanglement properties; second, to explore the feasibility of adiabatically driving the system between these configurations to implement a SWAP-like two-qubit gate operation.

\begin{quote}
    \centering
    \textit{“If you want to make a simulation of Nature, you'd better make it quantum mechanical."}\\
    \vspace{0.2cm}
    --- Richard P. Feynman    
\end{quote}



We structure our findings into three main sections. First, we undertake a general preliminary study of the system to verify key foundational properties. We confirm the negligible impact of the exchange interaction between wells at moderate inter-well separations, justifying our treatment of the system as two separable subsystems. Furthermore, we validate our numerical framework—specifically the Sinc discrete variable representation (Sinc-DVR)—by direct comparison with the known analytical solutions of the Morse potential, and we benchmark our proposed numerical methods for time-evolution using a simplified Landau-Zener model \eqref{eq:landau_zener}.

Subsequently, we report on the results of our parameter optimization routines, presenting optimal paramters for the two configurations $C_I$ and $C_{II}$, that achieve desired entanglement properties and targeted eigenstate population, illustrated visually by population plots.

Finally, we detail our time-dependent simulations. Here, we first demonstrate dynamics within a simplified two-level model, showcasing the feasibility of adiabatically driving the system between configurations $C_I$ and $C_{II}$ to implement desired population transfer. Following this, we present a more comprehensive time-evolution simulation of the full (four-level) two-particle Morse double-well system. This includes rigorous assessment of gate fidelity, and results incorporating phase correction techniques to enhance gate performance.

\newpage
\subfile{../sections/chap3/general_study_results.tex}
\subfile{../sections/chap3/Optimization_config.tex}
\subfile{../sections/chap3/time_evolution_res.tex}
\end{document}